\package{}: an \R{} package for implemetation of the ITHIM model

### 2016-09-15 manuscript/manuscript.tex
Worked on figure and changed permissions
I changed the permissins only so it looks nicer when using ls
Updates to ITHIM manuscript 1

\newcommand{\col}[2][red]{\textcolor{#1}{#2}}
\newcommand{\bi}{\begin{itemize}}
\newcommand{\ei}{\end{itemize}}
\newcommand{\package}{\emph{ITHIM}}
\newcommand{\ithim}{\emph{ITHIM}}
\newcommand{\R}{\emph{R}}
\newcommand{\sgy}[1]{{\itshape\col{#1}}}
\newcommand{\XX}{\sgy{XX}}
\newcommand{\webpage}{\url{http://www.ithim.ghi.wisc.edu}}

\begin{document}
\firstpage{1}

@@

\section{Summary:}
\col[blue]{While increased active transportation, primarily cycling and walk, is often held up as a goal for improved public health — via reduced chronic disease, improved air quality, and reductions in greenhouse gas emissions — obtaining estimates for the quantiative impacts of such shifts remains infrequent, in not altogether absent. Using U.S. metropolitan counties from  the 10 regions defined by
the Department of Health and Human Services (HHS1-10) we perform a
comparative risk analysis between each regions and national
averages as the baseline.  We investigated the reduced chronic disease burden of (x,y,z outcomes) from increased physical activity from population shifts in the amount of biking and walking doen for transport. 
Our methods employ the Integrated Transport and Health Impacts Model (ITHIM) informed by regional travel parameters
using the National Household Transportation Survey, non-travel activity metircs from the American Time Use Survey, and baseline Disease burden from the Centers for Disease Control and Prevention's WONDER database.  Our findings show that the West coast
and the Northeast of the U.S. have travel patterns that, if mirrored
by the nation, would save the U.S. thousands to hundreds of thousands
DALYs.}

@@

\section{Introduction}
\col{}

@@

\col[blue]{\begin{itemize}
  \item The coefficient of variation was fixed at ??? for all strata
    and models.  Earlier ITHIM models estimated this as a function of
    total active tranpsort time to model the decrease in variance of
    active transport time that occurs for increasing means, i.e.,
    fitter communities have less variance in active transport time.
  \item The mean non-travel activity for the referent group was
    modeled as a variable.  The relative means across age-sex strata
    were fixed, and estimated using ATUS.
  \item The national DALYs were stimated using the attributable
    fraction provided by the comparitive risk assesment.  The
    estimates for disease burned came from WHO and their GBD study.
  \item Not all of the regions had the same sample size.  In fact a
    few of the regions have relativel low sample size.  Maybe we
    should put confidence intervals on our estimates.
  \item The ITHIM model stratifies the population by sex and age using
    eight age classes.  Within each stratum active transport time is
    modeled as log-normal with mean $\hat{\mu}_{i,j}$ which is
    estimated from the National Household Transportation Survey.
  \item Quintiles, in this case the quantiles for percentages
    $(0.1,0.3,0.5,0.7,0.9)$, are used to discretize the integral
    defined in the Comparative Risk Assesment (CRA)
  \item 
\end{itemize}}

The ITHIM model uses the comparative risk assesment framework

@@

\col[red]{\begin{itemize}
\item National Household Transportation Survey (NHTS)
\item American time Use Survey
\item CDC Wonder
\end{itemize}}

@@

\col[red]{Lorem ipsum dolor sit amet, consectetur adipisicing elit, sed do
eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim ad
minim veniam, quis nostrud exercitation ullamco laboris nisi ut
aliquip ex ea commodo consequat. Duis aute irure dolor in
reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla
pariatur. Excepteur sint occaecat cupidatat non proident, sunt in
culpa qui officia deserunt mollit anim id est laborum. Lorem ipsum
dolor sit amet, consectetur adipisicing elit, sed do eiusmod tempor
incididunt ut labore et dolore magna aliqua. Ut enim ad minim veniam,
quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea
commodo consequat. Duis aute irure dolor in reprehenderit in voluptate
velit esse cillum dolore eu fugiat nulla pariatur. Excepteur sint
occaecat cupidatat non proident, sunt in culpa qui officia deserunt
mollit anim id est laborum. Lorem ipsum dolor sit amet, consectetur
adipisicing elit, sed do eiusmod tempor incididunt ut labore et dolore
magna aliqua. Ut enim ad minim veniam, quis nostrud exercitation
ullamco laboris nisi ut aliquip ex ea commodo consequat.}

@@

\begin{figure}[t]
    \centerline{\includegraphics[width=0.4\textwidth]{./figures/hhsmap}}
    \centerline{\includegraphics[width=0.4\textwidth]{./figures/fig2}}
    \caption{Change in total national DALYs for each regional
      transportation estimates.}\label{dalyFigure}
\end{figure}

\col[blue]{Lorem ipsum dolor sit amet, consectetur adipisicing elit, sed do
eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim ad
minim veniam, quis nostrud exercitation ullamco laboris nisi ut
aliquip ex ea commodo consequat. Duis aute irure dolor in
reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla
pariatur. Excepteur sint occaecat cupidatat non proident, sunt in
culpa qui officia deserunt mollit anim id est laborum. Lorem ipsum
dolor sit amet, consectetur adipisicing elit, sed do eiusmod tempor
incididunt ut labore et dolore magna aliqua. Ut enim ad minim veniam,
quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea
commodo consequat. Duis aute irure dolor in reprehenderit in voluptate
velit esse cillum dolore eu fugiat nulla pariatur. Excepteur sint
occaecat cupidatat non proident, sunt in culpa qui officia deserunt
mollit anim id est laborum. Lorem ipsum dolor sit amet, consectetur
adipisicing elit, sed do eiusmod tempor incididunt ut labore et dolore
magna aliqua. Ut enim ad minim veniam, quis nostrud exercitation
ullamco laboris nisi ut aliquip ex ea commodo consequat.}

@@

>>>>

\newcommand{\col}[2][red]{\textcolor{#1}{#2}}
\newcommand{\bi}{\begin{itemize}}
\newcommand{\ei}{\end{itemize}}
\newcommand{\package}{\emph{ITHIM}}
\newcommand{\ithim}{\emph{ITHIM}}
\newcommand{\R}{\emph{R}}
\newcommand{\sgy}[1]{{\itshape\col{#1}}}
\newcommand{\XX}{\sgy{XX}}
\newcommand{\webpage}{\url{http://www.ithim.ghi.wisc.edu}}
\newcommand{\af}{\rho}
\newcommand{\mets}{MET-hrs./week}
\begin{document}
\firstpage{1}

@@

\section{Summary:}
\col[blue]{While increased active transportation, primarily cycling
  and walk, is often held up as a goal for improved public health —
  via reduced chronic disease, improved air quality, and reductions in
  greenhouse gas emissions --- obtaining estimates for the quantiative
  impacts of such shifts remains infrequent, in not altogether
  absent. Using U.S. metropolitan counties from the 10 regions defined
  by the Department of Health and Human Services (HHS1-10) we perform
  a comparative risk analysis between each regions and national
  averages as the baseline.  We investigated the reduced chronic
  disease burden of (x,y,z outcomes) from increased physical activity
  from population shifts in the amount of biking and walking doen for
  transport.  Our methods employ the Integrated Transport and Health
  Impacts Model (ITHIM) informed by regional travel parameters using
  the National Household Transportation Survey, non-travel activity
  metircs from the American Time Use Survey, and baseline Disease
  burden from the Centers for Disease Control and Prevention's WONDER
  database.  Our findings show that the West coast and the Northeast
  of the U.S. have travel patterns that, if mirrored by the nation,
  would save the U.S. thousands to hundreds of thousands DALYs.}

@@

\section{Introduction}

@@

The ITHIM model uses the comparative risk assesment framework to
estimate the population attributable fraction, $\af$.  In our case
$\af$ represents the proportional change in mortality and morbidity if
the physical activity levels were adjusted to represent an alternate
scenario.

\begin{equation}
\af = \frac{\int R(x)P(x) dx  - \int R(x)Q(x) dx }{\int R(x)P(x) dx}
\end{equation}

Here $R(x)$ represents the risk when the individual has exposure $x$,
in our case physical activity (\mets).  $P(x)$ and $Q(x)$ denote the
population density of individuals with exposure $x$, in the baseline
and alternate scenarios, respectively.  The exposure variable, $x$, is
modeled as the sum of travel-related exposure and non-travel-related
exposure.

The time spent walking or cycling, i.e., active transport time, is
modeled as a log-normal random variable with means defined for each
age-sex stratum.  The non-travel-related physical activity is also
modeled as log-normal with stratum-specific means.  Estimates for
non-travel-related physical activity are diffiuclt.  We used the
American Time Use Survey data to estimate the relative means for each
of the age-sex strata.  We then treat the mean of the referent group
as a variable and compute results over a grid of values for this
variable.  See Figure \ref{dalyFigure}.

The travel-related exposure is estimated using active transport time
along with assumptions about how much physical activity is required
for cycling and walking.  Active transport time and non-travel-related
physical activity are modeled as independent random variables.  A
fixed value was used as the coefficient of variation for the
distributin of active transport time. A different, but also fixed,
value was used for the non-travel-related physical activity
distribution.

@@

@@

@@

\begin{figure}[t]
    \centerline{\includegraphics[width=0.4\textwidth]{./figures/HHSmap}}
    \centerline{\includegraphics[width=0.4\textwidth]{./figures/fig2}}
    \caption{Change in total national DALYs for each regional
      transportation estimates.}\label{dalyFigure}
\end{figure}

@@

%   \begin{itemize}
%   \item The coefficient of variation was fixed at ??? for all strata
% and models.  Earlier ITHIM models estimated this as a function of
% total active tranpsort time to model the decrease in variance of
% active transport time that occurs for increasing means, i.e., fitter
% communities have less variance in active transport time.
%   \item The mean non-travel activity for the referent group was
% modeled as a variable.  The relative means across age-sex strata were
% fixed, and estimated using ATUS.
%   \item The national DALYs were stimated using the attributable
% fraction provided by the comparitive risk assesment.  The estimates
% for disease burned came from WHO and their GBD study.
%   \item Not all of the regions had the same sample size.  In fact a
% few of the regions have relativel low sample size.  Maybe we should
% put confidence intervals on our estimates.
%   \item The ITHIM model stratifies the population by sex and age using
% eight age classes.  Within each stratum active transport time is
% modeled as log-normal with mean $\hat{\mu}_{i,j}$ which is estimated
% from the National Household Transportation Survey.
%   \item Quintiles, in this case the quantiles for percentages
% $(0.1,0.3,0.5,0.7,0.9)$, are used to discretize the integral defined
% in the Comparative Risk Assesment (CRA)
%   \item
% \end{itemize}
%  ``the proportional reduction in population
% disease or mortality that would occur if exposure to a risk factor
% were reduced to an alternative ideal exposure scenario (eg. no tobacco
% use).'' \url{http://www.who.int/healthinfo/global_burden_disease/metrics_paf/en/}
